```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(dplyr)
library(viridis)
library(purrr)
library(reshape2)
opts_chunk$set(echo = TRUE)
```

```{r}
# read yearly listing summary data
fnames = list.files('/Users/kang/Documents/syracuse_datascience/IST/719/proj/poster/02_presentation/data', pattern = 'listings_201\\d12.csv', full.names = TRUE)
df_list = lapply(
  fnames,
  function(x) fread(x)
)
# name each data set
names(df_list) <- 2015:2019

df <- map_df(df_list, ~as.data.frame(.x), .id="year")
dfplot <- df %>% count(year, neighbourhood_group)
```

### Plot: Growth rate in AirBnb listings in NYC has significantly decreased from 2015 - 2019

```{r}
nycGrowth <- data.frame(
  year = 2016:2019,
  perc_growth = round(diff(tapply(dfplot$n, dfplot$year, sum)) / tapply(dfplot$n, dfplot$year, sum)[1:4],2)
)

ggplot(nycGrowth, aes( x = year, y = perc_growth )) +
  geom_point(col='#F16664') + geom_line(col='#F16664') +
  ylab("percent change in listings") +
  ggtitle("Rate of Change in Number of Listings in NYC (2015 - 2019)")
```

### Plot: Growth in Listings Over Time

```{r}
dfplot$color <- '#6F5191'
dfplot$color[dfplot$neighbourhood_group=='Brooklyn'] <- '#7E689C'
dfplot$color[dfplot$neighbourhood_group=='Manhattan'] <- '#7F86A'
dfplot$color[dfplot$neighbourhood_group=='Queens'] <- '#526182'
dfplot$color[dfplot$neighbourhood_group=='Staten Island'] <- '#40456E'
ggplot(dfplot) +
  geom_col(aes(x=year, y=n, fill=neighbourhood_group)) +
  scale_fill_viridis(discrete = TRUE) +
  ylab("Count of Listings") +
  ggtitle("Total Number of AirBnb Listings in NYC by Year",
          subtitle = "Segmented by Neighbourhood Group")
```

### Plot: Overall trend shows an increase in listings, with the majority share of listings in Brooklyn and Manhattan.*

```{r}
# listings growth rate by neighbourhood
ngroups <- c('Bronx','Brooklyn','Manhattan','Queens','Staten Island')
# function to get YoY growth rate given neighbourhood group
getGrowth <- function(x) {
  d <- dfplot[dfplot$neighbourhood_group == x,]
  arr <- tapply(d$n, d$year, sum)
  out <- round(diff(arr)/arr[1:length(arr)-1], 2)
  return(out)
}
# dataframe of YoY change
dYearDelta <- as.data.frame(sapply(ngroups, getGrowth) %>% t)
dYearDelta$neighbour_group <- ngroups
dYearDelta <- dYearDelta[,c('neighbour_group','2016','2017','2018','2019')]
ggplot(melt(dYearDelta), aes(x = neighbour_group, y = variable, fill = value)) +
  geom_tile() +
  scale_fill_viridis() +
  ggtitle("Year-over-Year Growth in Listings by Neighbourhood Group")
```

### Plot: Distribution of Listing Price Over Time by Neighbourhood Group

```{r}
dfplot3 <- df[,c('year', 'neighbourhood_group', 'price')]
dfplot3$year <- as.factor(dfplot3$year)
# log transform price
dfplot3$logprice <- log(dfplot3$price)
ggplot(dfplot3, aes(y = logprice, factor(year), fill = year)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE) +
  xlab('year') +
  theme_dark() +
  ylab('log of listing price') +
  facet_wrap(~neighbourhood_group) +
  ggtitle("Log Listing Price by Year",
          subtitle = "Facets Neighbourhood Group")
```

### Plot: How Does Neighbourhood Persona (PRIZM Attributes)  Relate with Price?

```{r}
# cleaned listings data for 2019, including census attributes
df <- fread('/Users/kang/Documents/syracuse_datascience/IST/719/proj/poster/02_presentation/data/listings_cleaned.csv')
```

```{r}
# discretize price into 5 equal frequency bins
df$priceDisc <- arules::discretize(df$price, method = 'frequency', categories = 5)
```

```{r}
dat <- as.data.frame(df)
# remove prizm_cluster = 'UNK' (unknown)
dat <- dat[!dat$prizm_cluster == 'UNK',]
# keep just the top prizm clusters that make up 80% of the data
clust_freq <- sort(table(dat$prizm_cluster), decreasing = TRUE)
clust_keep <- clust_freq[which(cumsum(clust_freq) <= 0.8 * sum(clust_freq))] %>% names
dat <- dat[dat$prizm_cluster %in% clust_keep,]
# normalize the prizm features for each cluster
# function to normalize vector
normPrizm <- function(x) {
  paste(
    unique(
      do.call('c', strsplit(gsub("\\[|\\]|[\"']",'',x), split = ', '))
    ), collapse = ' | '
  )
}
for (i in clust_keep) {
  dat$prizm_features_norm[dat$prizm_cluster == i] <- normPrizm(
    dat$prizm_features[dat$prizm_cluster == i]
  )
}
# contingency table of clusters and discretized price
dfPlot <- table(dat[,'prizm_cluster'], dat[,'priceDisc']) %>% as.data.frame
# reorder cluster factor levels
dfPlot$Var1 <- factor(dfPlot$Var1, levels = clust_keep)
# scale frequency relative to cluster
for (i in clust_keep) {
  dfPlot$Freq[dfPlot$Var1 == i] <- scale(dfPlot$Freq[dfPlot$Var1 == i])
}
dfPlot <- transform(dfPlot, Var1=as.numeric(Var1))
```

```{r}
ggplot(dfPlot, aes(Var1, Var2)) + 
  geom_point(size = 10, aes(color = Freq)) + 
  scale_color_viridis() +
  xlab("Cluster Label") + ylab("Price Bins") +
  ggtitle("Contingency Plot of Discretized Price and Top 14 Neighbourhood Personas")
```

### Plot: How does the number of listings by host relate to price; Segmented by Neighbourhood Group

```{r}
df$logprice <- log(df$price)
dfplot <- df[,c('logprice','host_listings_count','neighbourhood_group_cleansed')]
colnames(dfplot) <- c('logprice','host_listings_count','neighbourhood_grp')

ggplot(dfplot, aes(x = host_listings_count, y = logprice, color = neighbourhood_grp)) +
  facet_wrap(~neighbourhood_grp, scales = 'free_x') +
  geom_point() +
  ggtitle("Host Listings Count vs. Log Price")
```

### Plot: map

```{r}
# transform data for plot
# get latitude and longitude coordinates from original data
keep_cols <- c(
  'id','state','price','population','median_home_value',
  'median_household_income','neighbourhood_group_cleansed'
)
# listings_cleaned
df <- as.data.frame(df)[,keep_cols] 
# original 2019 listings summary
df2 <- as.data.frame(df_list[[5]])[,c('id','latitude','longitude')]  
dfplot <- merge(df, df2, by = 'id') %>% .[complete.cases(.),]
# filter on just state == 'NY'
dfplot <- dfplot[dfplot$state == 'NY',]
# change 'NY' to 'new york'
dfplot$state <- 'new york'
```

```{r}
# specify latitude longitude coordinate limits
xlim.min = min(dfplot$longitude)
xlim.max = max(dfplot$longitude)
ylim.min = min(dfplot$latitude)
ylim.max = max(dfplot$latitude)
```

```{r}
# specify state map data
us <- map_data('state')
```

### Density of Listings

```{r}
# Density of listings
ggplot(dfplot, aes(map_id = state, x = longitude, y = latitude)) +
  geom_map(map = us, fill = 'black', color = 'white') +
  expand_limits(x = us$long, y = us$lat) +
  coord_map() +
  scale_color_viridis() +
  theme_bw() +
  stat_density2d(aes(color = ..level..)) +
  xlim(xlim.min, xlim.max) + ylim(ylim.min, ylim.max) +
  ggtitle("Density of Listings in New York")
```

```{r}
# median home value income by Zipcode
ggplot(dfplot, aes(map_id = state)) +
  geom_map(map = us, fill='white', color = 'black') +
  expand_limits(x = us$long, y = us$lat) +
  coord_map() +
  geom_point(aes(x = longitude, y = latitude, col = median_household_income)) +
  scale_color_viridis() +
  theme_bw() +
  xlim(xlim.min, xlim.max) + ylim(ylim.min, ylim.max) +
  ggtitle("Median Income by Zipcode in New York")
```
